### 2016-01-04 conclusions.tex
Minor fixes
minor changes to the text

We have investigated scale-free algorithms for online linear optimization and
we have shown that scale-free property leads to algorithms which have optimal
regret and do not need to know or assume \textbf{anything} about the sequence
of loss vectors. In particular, the algorithms do not assume any upper or lower
bounds on the norms of the loss vectors or the number of rounds.

@@

We have shown an $\frac{D}{\sqrt{8}} \sqrt{\sum_{t=1}^T \|\ell\|_*^2}$
lower bound on regret of any algorithm for decision set with diameter $D$.

We have shown that one of the scale-free \textsc{FTRL}-based the algorithms has
regret $O \left(f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} + \max_{t=1,2,\dots,T}
\|\ell_t\|_* \sqrt{T} \right)$ with respect to any competitor $u \in K$, where
$f$ is any non-negative $1$-strongly convex function defined on the decision
set.  The latter result makes sense even when the decision set $K$ is
unbounded. Notice that with regularizer $f(u) = \frac{1}{2}\|u\|_2^2$ the
regret depends on norm of the competitor $\|u\|_2$ is quadratically. There
exist non-scale-free algorithms \cite{McMahan-Streeter-2012,
McMahan-Abernethy-2013, Orabona-2013, McMahan-Orabona-2014, Orabona-2014} that
have only $O(\|u\|_2 \sqrt{\log \|u_2\|_2})$ or $O(\|u\|_2 \log \|u\|_2)$
dependency.  These algorithms assume an a priori bound on the norm of the loss
vectors.  This leads to another open problem:
%
\begin{quotation}
\noindent
\emph{Is there a scale-free algorithm for the decision set $\R^d$, regret of
which has $O(\|u\|_2 \sqrt{\log \|u\|_2})$ dependency on the competitor $u$?}
\end{quotation}

>>>>

We have investigated scale-free algorithms for online linear optimization and
we have shown that the scale-free property leads to algorithms which have optimal
regret and do not need to know or assume \textbf{anything} about the sequence
of loss vectors. In particular, the algorithms do not assume any upper or lower
bounds on the norms of the loss vectors or the number of rounds.

@@

We have proved an $\frac{D}{\sqrt{8}} \sqrt{\sum_{t=1}^T \|\ell\|_*^2}$
lower bound on regret of any algorithm for decision set with diameter $D$.

We have shown that one of the scale-free \textsc{FTRL}-based the algorithms has
regret $O \left(f(u) \sqrt{\sum_{t=1}^T \|\ell_t\|_*^2} + \max_{t=1,2,\dots,T}
\|\ell_t\|_* \sqrt{T} \right)$ with respect to any competitor $u \in K$, where
$f$ is any non-negative $1$-strongly convex function defined on the decision
set.  The latter result makes sense even when the decision set $K$ is
unbounded. Notice that with the regularizer $f(u) = \frac{1}{2}\|u\|_2^2$ the
regret depends quadratically on norm of the competitor $\|u\|_2$. There
exist non-scale-free algorithms \cite{McMahan-Streeter-2012,
McMahan-Abernethy-2013, Orabona-2013, McMahan-Orabona-2014, Orabona-2014} that
have only $O(\|u\|_2 \sqrt{\log \|u_2\|_2})$ or $O(\|u\|_2 \log \|u\|_2)$
dependency.  These algorithms assume an a priori bound on the norm of the loss
vectors.  This leads to another open problem:
%
\begin{quotation}
\noindent
\emph{Is there a scale-free algorithm for the decision set $\R^d$, regret of
which has $O(\|u\|_2 \sqrt{\log \|u\|_2})$ dependency on the competitor $u$?}
\end{quotation}

### 2016-01-04 mirror-descent.tex
Minor fixes
minor changes to the text

\textsc{Mirror Descent} (MD) is a generic algorithm similar to \textsc{FTRL}
but quite different in details. The algorithm is stated as
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
parametrized by a sequence $\{R_t\}_{t=0}^\infty$ of convex functions $R_t:K
\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on past
loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an arbitrary way. If $R_t$ is
not differentiable\footnote{Note that this can happen even when $R_t$ is a
restriction of a differentiable function defined on a superset of $K$.  If $K$
is bounded and closed, $R_t$ fails to be differentiable at the boundary of $K$.
If $K$ is a subset of an affine subspace of a dimension smaller than the
dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, the
Bregman divergence, $\Breg_{R_t}(u,v) = R_t(u) - R_t(v) - \langle \grad R_t(v),
u - v \rangle$ needs to be defined. This is done by choosing a subgradient map
$\grad R_t:K \to V$, i.e. a function such that $\grad R_t(w)$ is a subgradient
of $R_t$ at a point $w$. If $R_t$ is a restriction of a differentiable function
$R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
\in K$. The following lemma bounds regret of \textsc{MD}.  Proof can be found
in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.  For
completeness we give a proof in~\ref{section:mirror-descent-proofs}.

\begin{lemma}[Regret of Mirror Descent]
\label{lemma:mirror-descent-regret}
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer} satisfies, for any
$u \in K$,
$$
\Regret_T(u)
\le
\sum_{t=1}^T \langle \ell_t, w_t - w_{t+1} \rangle - \Breg_{R_t}(w_{t+1}, w_t) + \Breg_{R_t}(u,w_t) - \Breg_{R_t}(u, w_{t+1}) \; .
$$
\end{lemma}

@@

The most prominent instance of \textsc{MD} is \textsc{Projected Gradient
Descent} defined by the recurrence
$$
w_{t+1} = \Pi_K \left( w_t - \eta_t \ell_t \right)
$$
where $\Pi_K(x) = \argmin_{y \in K} \|x-y\|_2$ is the Euclidean projection to
$K$ and $\eta_t > 0$ is the step size.  \textsc{Projected Gradient Descent} is
an instance of \textsc{MD} with regularizer $R_t(w) =
\frac{1}{2\eta_t}\|w\|_2^2$.  A special case of \textsc{Projected Gradient
Descent} with step size $\eta_t = \frac{1}{\sqrt{t}}$ is \textsc{Generalized
Infinitesimal Gradient Ascent} (GIGA)~\cite{Zinkevich-2003}.  Full-matrix and
diagonal versions of \textsc{AdaGrad with Composite MD
Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of \textsc{MD} for any
bounded $K$ with regularizers
$$
R_t(w) = \frac{1}{2} w^\top \left(\sum_{i=1}^t \ell_i \ell_i^\top \right)^{1/2} \!\!\!\! w
\qquad \text{and} \qquad
R_t(w) = \frac{1}{2} \sum_{j=1}^d w_j^2 \sqrt{ \sum_{i=1}^t \ell_{i,j}^2}
$$
respectively. The diagonal version is simply the per-coordinate construction
applied to the one-dimensional version. When $d=1$, both algorithms are
equivalent to
\begin{equation}
\label{equation:ada-grad-regularizer}
R_t(w) = \frac{\|w\|_2^2}{2} \sqrt{\sum_{i=1}^t \|\ell_i\|_2^2 } \; .
\end{equation}
All three sequences of regularizers give a scale-free algorithm. Regularizer
\eqref{equation:ada-grad-regularizer} can be viewed as a scale-free
generalization of GIGA. That is, when $\|\ell_i\|_2
= 1$ for all $i$, \textsc{MD} with regularizer
\eqref{equation:ada-grad-regularizer} coincides with GIGA.

@@

The regret bound for \textsc{Scale-Free MD} in the
Corollary~\ref{corollary:regret-scale-free-mirror-descent} depends on
$\sup_{u,v \in K} \Breg_f(u,v)$. In contrast, the regret bounds for
\textsc{AdaFTRL} and \textsc{SOLO FTRL} in
Corollaries~\ref{corollary:ada-ftrl-regret-bound}~and~\ref{corollary:regret-solo-ftrl-bounded-set}
depend on $\sup_{u \in K} f(u)$.  Similarly, the regret bound in
Theorem~\ref{theorem:regret-scale-free-mirror-descent} for \textsc{Scale-Free
MD} depends on $\sup_{v \in K} \Breg_R(u,v)$ and the regret bounds in
Theorems~\ref{theorem:ada-ftrl-regret-bound}~and~\ref{theorem:regret-solo-ftrl}
for \textsc{AdaFTRL} and \textsc{SOLO FTRL} depend on $R(u)$. It is not hard to
show that
\begin{equation}
\label{equation:regularizer-vs-divergence}
\forall u \in K \qquad R(u) \le \sup_{v \in K} \Breg_R(u,v)
\end{equation}
provided that at the minimizer $v^* = \argmin_{v \in K} R(v)$ both $R(v^*)$ and
$\grad R(v^*)$ are zero. Indeed, in that case, $R(u) = \Breg_R(u,v^*) \le
\sup_{v \in K} \Breg_R(u,v)$.

>>>>

\textsc{Mirror Descent} (MD) is a generic algorithm similar to \textsc{FTRL}
but quite different in details. The algorithm is stated as
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
parametrized by a sequence $\{R_t\}_{t=0}^\infty$ of convex functions $R_t:K
\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on past
loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an arbitrary way. If $R_t$ is
not differentiable\footnote{Note that this can happen even when $R_t$ is a
restriction of a differentiable function defined on a superset of $K$.  If $K$
is bounded and closed, $R_t$ fails to be differentiable at the boundary of $K$.
If $K$ is a subset of an affine subspace of a dimension smaller than the
dimension of $V$, then $R_t$ fails to be differentiable everywhere.}, the
Bregman divergence, $\Breg_{R_t}(u,v) = R_t(u) - R_t(v) - \langle \grad R_t(v),
u - v \rangle$ needs to be defined. This is done by choosing a subgradient map
$\grad R_t:K \to V$, i.e., a function such that $\grad R_t(w)$ is a subgradient
of $R_t$ at a point $w$. If $R_t$ is a restriction of a differentiable function
$R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
\in K$. The following lemma bounds regret of \textsc{MD}.  Proof can be found
in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.  For
completeness we give a proof in~\ref{section:mirror-descent-proofs}.

\begin{lemma}[Regret of \textsc{MD}]
\label{lemma:mirror-descent-regret}
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer} satisfies, for any
$u \in K$,
$$
\Regret_T(u)
\le
\sum_{t=1}^T \langle \ell_t, w_t - w_{t+1} \rangle - \Breg_{R_t}(w_{t+1}, w_t) + \Breg_{R_t}(u,w_t) - \Breg_{R_t}(u, w_{t+1}) \; .
$$
\end{lemma}

@@

The most prominent instance of \textsc{MD} is \textsc{Projected Gradient
Descent} defined by the recurrence
$$
w_{t+1} = \Pi_K \left( w_t - \eta_t \ell_t \right),
$$
where $\Pi_K(x) = \argmin_{y \in K} \|x-y\|_2$ is the Euclidean projection to
$K$ and $\eta_t > 0$ is the step size.  \textsc{Projected Gradient Descent} is
an instance of \textsc{MD} with regularizer $R_t(w) =
\frac{1}{2\eta_t}\|w\|_2^2$.  A special case of \textsc{Projected Gradient
Descent} with step size $\eta_t = \frac{1}{\sqrt{t}}$ is \textsc{Generalized
Infinitesimal Gradient Ascent} (GIGA)~\cite{Zinkevich-2003}.  Full-matrix and
diagonal versions of \textsc{AdaGrad with Composite MD
Update}~\cite{Duchi-Hazan-Singer-2011} are an instances of \textsc{MD} for any
bounded $K$ with regularizers
$$
R_t(w) = \frac{1}{2} w^\top \left(\sum_{i=1}^t \ell_i \ell_i^\top \right)^{1/2} \!\!\!\! w
\qquad \text{and} \qquad
R_t(w) = \frac{1}{2} \sum_{j=1}^d w_j^2 \sqrt{ \sum_{i=1}^t \ell_{i,j}^2}
$$
respectively. The diagonal version is simply the per-coordinate construction
applied to the one-dimensional version. When $d=1$, both algorithms are
equivalent to
\begin{equation}
\label{equation:ada-grad-regularizer}
R_t(w) = \frac{\|w\|_2^2}{2} \sqrt{\sum_{i=1}^t \|\ell_i\|_2^2 } \; .
\end{equation}
All three sequences of regularizers give a scale-free algorithm. Regularizer
\eqref{equation:ada-grad-regularizer} can be viewed as a scale-free
generalization of GIGA. That is, when $\|\ell_i\|_2
= 1$ for all $i$, \textsc{MD} with regularizer
\eqref{equation:ada-grad-regularizer} coincides with GIGA.

@@

The regret bound for \textsc{Scale-Free MD} in the
Corollary~\ref{corollary:regret-scale-free-mirror-descent} depends on
$\sup_{u,v \in K} \Breg_f(u,v)$. In contrast, the regret bounds for
\textsc{AdaFTRL} and \textsc{SOLO FTRL} in
Corollaries~\ref{corollary:ada-ftrl-regret-bound}~and~\ref{corollary:regret-solo-ftrl-bounded-set}
depend on $\sup_{u \in K} f(u)$.  Similarly, the regret bound in
Theorem~\ref{theorem:regret-scale-free-mirror-descent} for \textsc{Scale-Free
MD} depends on $\sup_{v \in K} \Breg_R(u,v)$ and the regret bounds in
Theorems~\ref{theorem:ada-ftrl-regret-bound}~and~\ref{theorem:regret-solo-ftrl}
for \textsc{AdaFTRL} and \textsc{SOLO FTRL} depend on $R(u)$. It is not hard to
show that
\begin{equation}
\label{equation:regularizer-vs-divergence}
\forall u \in K \qquad R(u) \le \sup_{v \in K} \Breg_R(u,v),
\end{equation}
provided that at the minimizer $v^* = \argmin_{v \in K} R(v)$ both $R(v^*)$ and
$\grad R(v^*)$ are zero. Indeed, in that case, $R(u) = \Breg_R(u,v^*) \le
\sup_{v \in K} \Breg_R(u,v)$.

### 2016-01-04 preliminaries.tex
minor changes to the text

Monotonicity of Fenchel conjugates follows easily from the definition: If
$f,g:K \to \R$ satisfy $f(w) \le g(w)$ for all $w \in K$ then $f^*(\ell) \ge
g^*(\ell)$ for every $\ell \in V^*$.

@@

\begin{lemma}[Regret of Follow The Regularized Leader]
\label{lemma:generic-regret-bound}
If the regularizers $R_1, R_2, \dots$ chosen by
Algorithm~\ref{algorithm:ftrl-varying-regularizer} are strongly convex and lower
semi-continuous, the algorithm's regret is upper bounded as
$$
\Regret_T(u) \le R_{T+1}(u) + R_1^*(0) + \sum_{t=1}^{T} \Breg_{R_t^*}(-L_t, -L_{t-1}) - R_t^*(-L_t) + R_{t+1}^*(-L_t) \; .
$$
\end{lemma}

>>>>

Monotonicity of Fenchel conjugates follows easily from the definition: If
$f,g:K \to \R$ satisfies $f(w) \le g(w)$ for all $w \in K$ then $f^*(\ell) \ge
g^*(\ell)$ for every $\ell \in V^*$.

@@

\begin{lemma}[Regret of \textsc{FTRL}]
\label{lemma:generic-regret-bound}
If the regularizers $R_1, R_2, \dots$ chosen by
Algorithm~\ref{algorithm:ftrl-varying-regularizer} are strongly convex and lower
semi-continuous, the algorithm's regret is upper bounded as
$$
\Regret_T(u) \le R_{T+1}(u) + R_1^*(0) + \sum_{t=1}^{T} \Breg_{R_t^*}(-L_t, -L_{t-1}) - R_t^*(-L_t) + R_{t+1}^*(-L_t) \; .
$$
\end{lemma}
